Global patterns of forest autotrophic carbon fluxes

### 2020-07-25 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
mostly issue #79

Length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C. We experimented with a definion of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. However, we found that including a moisture index had minimal effect on the estimates of growing season length, and so chose to exclude it. Monthly data for PET, precipitation, and temperature from the CRU dataset v 4.03 [@harris_updated_2014], and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season. Total growing season precipitation and solar radiation were also calculated.

@@

When FACF were standardized by growing season length (in monthly increments), correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs. S8-S9). Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.

@@

FACF are negatively correlated with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014]. We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough **BBL: ?** with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow. A temperature-defined growing season length had stong positive correlation with FACF (Fig. 4f), but explained less variation than $MAT$. Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007]. However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season. We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power. Mechanistic accounting for climatic effects on global FACF patterns instead requires models representing physiologically meaningful timescales (*e.g.*, **refs**) [@longo_biophysics_2019]

Our analysis clarifies how FACF vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming. Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of FACF (Figs. 2-3). This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018]. **DISCUSS/CITE [Wu et al. 2020]** [@wu_accelerated_2020] However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances. **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)**
Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across geographic gradients in climate. Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

>>>>

**For consistency with previous studies (Table 1, *H5*),** length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C. We experimented with a definition of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. However, we found that including a moisture index had minimal effect on the estimates of growing season length, and so chose to exclude it. Monthly data for PET, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season. Total growing season precipitation and solar radiation were also calculated.

@@

When FACF were standardized by growing season length (in monthly increments), correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs. S8-S9). Specifically, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.

@@

Forest autotrophic C fluxes decline with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014]. The sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it's calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow. We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig. 4f), but explained less variation than $MAT$. Dividing annual fluxes by growing season length to yield average flux per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007]. However, there remained a number of significant correlations with growing season climatic conditions, indicating that climatic conditions remain influential within the growing season. We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power. Mechanistic accounting for climatic effects on global forest carbon flux patterns instead requires models representing physiologically meaningful timescales [e.g., **REFS**; @longo_biophysics_2019].

Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming. Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of FACF (Figs. 2-3). This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018]. **DISCUSS/CITE [Wu et al. 2020]** [@wu_accelerated_2020] However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances. **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)**
Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across geographic gradients in climate. Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
